### 2022-05-05 presentation/presentation.tex
last few updates

\section{Introduction}
\begin{frame}{Introduction}{Problem Setting}
	\begin{itemize}
		\item Observational studies frequent in econometrics, psychology, and medical research
		\item This presentation: binary treatment/control variable, $\vec{z}$.
		\item Goal is to estimate ATE: $\mathbbmss E [y_{i1} - y_{i0}]$
		\item Solve this problem via \emph{matching} treatment observations to control observations
		\item We break this down into two orthogonal problems:
			\begin{itemize}
				\item How is distance measured?
				\item How are matches assigned?
			\end{itemize}
		\item \texttt{matching}: Python package for matching observational data.
	\end{itemize}
\end{frame}

\begin{frame}{Introduction}{Notation}
	Now, we will introduce some notation, so we can precisely talk about construction of different matching methods.
	\begin{itemize}
		\item $\mat{X} \in \reals^{(n + m) \times p}$ of features; $\mat{X}_T \in \reals^{n \times p}, \mat{X}_C \in \reals^{m \times p}$
		\item $\vec{z} \in \{0, 1\}^{n + m}$ of binary treatment assignments;
		\item $d: \reals^p \times \reals^p \to \reals^+$ is a distance measure between feature vectors, e.g. $d(\vec{x}_{T_{i}}, \vec{x}_{C_{j}})$;
		\item $\mathcal{D}_d : \reals^{n \times p} \times \reals^{m \times p} \to \reals^{n \times m}$ produces \emph{biadjacency matrix};
		\item Some procedures use a preprocessing function $f$ (usually: dimension reduction or coarsening of the data).
	\end{itemize}
	Our framework is similar to that of~\textcite{iacus_multivariate_2011}:
	\begin{equation}
		\mathcal{D}_d(f(\mat{X})_T, f(\mat{X})_C). \label{eqn:beamerabstractdistance}
	\end{equation}
\end{frame}

@@

\begin{frame}{Measuring Distance}{The Propensity Score}
	\begin{definition}[propensity score, \textcite{rosenbaum_central_1983}]
		The \emph{propensity score} is the balancing score \( b(\mat{X}) = \mathbbmss E [\vec{z} | \mat{X}] \).
	\end{definition}%
	\begin{itemize}
		\item Even if not used for matching, a useful diagnostic~\parencite{dehejia_causal_1999}.
		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{z}$.
		\item Continuous scores: \textcite{imai_causal_2004}.
		\item Often prune potential matches with a caliper $c$.
	\end{itemize}
\end{frame}

\begin{frame}{Measuring Distance}{The Prognostic Score}
	\begin{definition}[prognostic score, \textcite{hansen_prognostic_2008}]
		The \emph{prognostic score} is the function \( b(\mat{X}) = \mathbbmss E_{C} [\vec{y} | \mat{X} ] \).
	\end{definition}
	\begin{itemize}
		\item Note that the expectation is with respect to control data $C$.
		\begin{itemize}
			\item Previous attempts to match on outcomes failed, such as \textcite{miettinen_stratification_1976}.
		\end{itemize}
		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{y}$.
		\item  ``conditionality principle'' suggests usefulness \textcite{hansen_prognostic_2008}.
		\item Joint use of propensity and prognostic scores \textcite{leacy_joint_2014}.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Bipartite Matching Algorithms}{}
	Define a matching $\set{M}$ as a set of tuples $(i, j)$ indicating a matching between observations $T_i$ and $C_j$. 
	``Optimal'' matching via the Hungarian algorithm \parencite{munkres_algorithms_1957}.
	\begin{equation}
		\min _{\mathcal{M}} \sum_{(i, j) \in \mathcal{M}} d\left(\vec{x}_{T_{i}}, \vec{x}_{C_{j}}\right), \text { s.t. } \forall(i, j),(k, l) \in \mathcal{M}, i=k \iff j=l.
	\end{equation}
	\begin{itemize}
		\item Extendable to $1:k$ matching via $b$-matching.
		\item Slow: $O((n + m)^3)$; approximation: $b$-\textsc{Suitor}~\parencite{khan_efficient_2016}.
	\end{itemize}

	Greedy matching: sort edges by distance, pop matches off the top.
	\begin{itemize}
		\item No optimality assured.
		\item Problems when matching $1:k$ \parencite{rosenbaum_optimal_1989}.
	\end{itemize}
\end{frame}

@@

\section{Experiments}
\begin{frame}{Experiments}{Overview}
	\begin{itemize}
		\item Hyperparameters $n$, $m$, $p$, $\theta_0$, $\theta_1$, and $\rho$.
	\begin{align}
		\vec{\upmu}_0 \sim \mathcal N(\theta_0, 0.25), \vec{\upmu}_1 \sim \mathcal N(\theta_1, 0.25) \\
		\{\vec{x}_C\}_{j=1}^{m} \sim \mathcal N(\vec{\upmu}_0, \Sigma), \{\vec{x}_T\}_{i=1}^{n} \sim \mathcal N(\vec{\upmu}_1, \Sigma) \\
		\text{ where } \Sigma_{a b} = \begin{cases}
			1 & \text{ if } a = b \\
			\rho & \text{ otherwise.}
		\end{cases}
	\end{align}
		\item Balance metrics: mean ASMD\footnote{Absolute Standardized Mean Difference}, maximum ASMD, proportion of treatment observations matched.
		\begin{itemize}
			\item In practice: use more robust metrics \parencite{basu_use_2008, zhu_kernel-based_2018}.
		\end{itemize}
		\item Vary calipers $c \in \{0.05, 0.10, 0.15, \ldots, 0.50\}$.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Experiments}{Caliper vs. Imbalance: Mean Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_mean_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Imbalance: Maximum Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_calipermax_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Imbalance: Proportion of Treatment Observations Matched}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_prop1_match.png}
	\end{center}
\end{frame}

\subsection{Caliper vs. Correlation}
\begin{frame}{Experiments}{Caliper vs. Correlation}
	This will consider the effects of increased correlation between the features of $X$
	\begin{itemize}
		\item $m = 750$
		\item $n = 250$
		\item $p = 5$
		\item $\theta_0 = 0$
		\item $\theta_1 = 1$
	\end{itemize}
	We vary $\rho \in \{0, 0.1, 0.2, \ldots, 0.8\}$.
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Mean Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_mean_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Maximum Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_calipermax_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Proportion of Treatment Observations Matched}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_prop1_match.png}
	\end{center}
\end{frame}

>>>>

\section{Introduction}
\begin{frame}{Introduction}{Problem Setting}
	\begin{itemize}
		\item Observational studies frequent in econometrics, psychology, and medical research.
		\item This presentation: binary treatment/control variable, $\vec{z}$.
		\item Most often, goal is to estimate ATE: $\mathbbmss E [y_{i1} - y_{i0}]$.
		\item Solve this problem via \emph{matching} treatment observations to control observations
			\begin{itemize}
				\item Produce a matched subset suitable for analysis
			\end{itemize}
		\item We break this down into two orthogonal problems:
			\begin{itemize}
				\item How is distance measured?
				\item How are matches assigned?
			\end{itemize}
		\item \texttt{matching}: Python package for matching observational data.
	\end{itemize}
\end{frame}

\begin{frame}{Introduction}{Notation}
	Now, we will introduce some notation, so we can precisely talk about construction of different matching methods.
	\begin{itemize}
		\item $\mat{X} \in \reals^{(n + m) \times p}$ of features; $\mat{X}_T \in \reals^{n \times p}, \mat{X}_C \in \reals^{m \times p}$, $n \leq m$;
		\item $\vec{z} \in \{0, 1\}^{n + m}$ of binary treatment assignments;
		\item $d: \reals^p \times \reals^p \to \reals^+$ is a distance measure between feature vectors, e.g. $d(\vec{x}_{T_{i}}, \vec{x}_{C_{j}})$;
		\item $\mathcal{D}_d : \reals^{n \times p} \times \reals^{m \times p} \to \reals^{n \times m}$ produces \emph{biadjacency matrix};
		\item Some procedures use a preprocessing function $f$ (usually: dimension reduction or coarsening of the data).
	\end{itemize}
	Our framework is similar to that of~\textcite{iacus_multivariate_2011}:
	\begin{equation}
		\mathcal{D}_d(f(\mat{X})_T, f(\mat{X})_C). \label{eqn:beamerabstractdistance}
	\end{equation}
\end{frame}

@@

\begin{frame}{Measuring Distance}{The Propensity Score}
	\begin{definition}[propensity score, \textcite{rosenbaum_central_1983}]
		The \emph{propensity score} is the balancing score \( b(\mat{X}) = \mathbbmss E [\vec{z} | \mat{X}] \).
	\end{definition}%
	\begin{itemize}
		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{z}$.
		\item Often prune potential matches with a caliper $c$.
		\item Even if not used for matching, a useful diagnostic~\parencite{dehejia_causal_1999}.
		\item Continuous scores: \textcite{imai_causal_2004}.
	\end{itemize}
\end{frame}

\begin{frame}{Measuring Distance}{The Prognostic Score}
	\begin{definition}[prognostic score, \textcite{hansen_prognostic_2008}]
		The \emph{prognostic score} is the function \( b(\mat{X}) = \mathbbmss E_{C} [\vec{y} | \mat{X} ] \).
	\end{definition}
	\begin{itemize}
		\item Note that the expectation is with respect to control data $C$.
		\begin{itemize}
			\item No information about ATE
			\item Previous attempts to match on outcomes failed, such as \textcite{miettinen_stratification_1976}.
			\item  ``conditionality principle'' \textcite{hansen_prognostic_2008}.
		\end{itemize}
		\item Implicitly ``weights'' features by heterogeneity wrt $\vec{y}$.
	\end{itemize}
\end{frame}

@@

\subsection{Matching Algorithms}
\begin{frame}{Bipartite Matching Algorithms}{}
	Define a matching $\set{M}$ as a set of tuples $(i, j)$ indicating a matching between observations $T_i$ and $C_j$. 
	``Optimal'' matching via the Hungarian algorithm \parencite{munkres_algorithms_1957}.
	\begin{equation}
		\min _{\mathcal{M}} \sum_{(i, j) \in \mathcal{M}} d\left(\vec{x}_{T_{i}}, \vec{x}_{C_{j}}\right), \text { s.t. } \forall(i, j),(k, l) \in \mathcal{M}, i=k \iff j=l.
	\end{equation}
	\begin{itemize}
		\item Extendable to $1:k$ matching via $b$-matching.
		\item Slow: $O((k \cdot n + m)^3)$; approximation: $b$-\textsc{Suitor}~\parencite{khan_efficient_2016}.
	\end{itemize}

	Greedy matching: sort edges by distance, pop matches off the top.
	\begin{itemize}
		\item When control observations may not be reused, no optimality assured.
		\item Problems when matching $1:k$ \parencite{rosenbaum_optimal_1989}.
	\end{itemize}
\end{frame}

@@

\section{Experiments}
\begin{frame}{Experiments}{Overview}
	\begin{itemize}
		\item Hyperparameters $n$, $m$, $p$, $\theta_0$, $\theta_1$, and $\rho$.
	\begin{align}
		\vec{\upmu}_0 \sim \mathcal N(\theta_0, 0.25), \vec{\upmu}_1 \sim \mathcal N(\theta_1, 0.25) \\
		\{\vec{x}_C\}_{j=1}^{m} \sim \mathcal N(\vec{\upmu}_0, \Sigma), \{\vec{x}_T\}_{i=1}^{n} \sim \mathcal N(\vec{\upmu}_1, \Sigma) \\
		\text{ where } \Sigma_{a b} = \begin{cases}
			1 & \text{ if } a = b \\
			\rho & \text{ otherwise.}
		\end{cases}
	\end{align}
		\item Metrics: mean ASMD\footnote{Absolute Standardized Mean Difference}, proportion of treatment observations matched.
		\begin{itemize}
			\item In practice: use more robust metrics \parencite{basu_use_2008, zhu_kernel-based_2018}.
		\end{itemize}
		\item Vary calipers $c \in \{0.05, 0.10, 0.15, \ldots, 0.50\}$.
	\end{itemize}
\end{frame}

@@

\begin{frame}{Experiments}{Caliper vs. Imbalance: Mean Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_mean_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Imbalance: Proportion of Treatment Observations Matched}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_imbalance_big/plots/theta1_caliper_prop1_match.png}
	\end{center}
\end{frame}

\subsection{Caliper vs. Correlation}
\begin{frame}{Experiments}{Caliper vs. Correlation}
	This will consider the effects of increased correlation between the features of $X$
	\begin{itemize}
		\item $m = 750$
		\item $n = 250$
		\item $p = 5$
		\item $\theta_0 = 0$
		\item $\theta_1 = 1$
	\end{itemize}
	We vary $\rho \in \{0, 0.1, 0.2, \ldots, 0.8\}$.
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Mean Absolute Standardized Mean Difference}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_mean_abs_smd.png}
	\end{center}
\end{frame}
\begin{frame}{Experiments}{Caliper vs. Correlation: Proportion of Treatment Observations Matched}
	\begin{center}
		\includegraphics[width=\textwidth]{../paper/img/output30/caliper_vs_correlation_big/plots/rho_caliper_prop1_match.png}
	\end{center}
\end{frame}
